Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-25 EconGrowth.Rnw
Committed text changes.
Now drawing a graph of US GDP vs. time data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Caleb, put your LaTeX code here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<setup, include=FALSE>>=
require(knitr)
require(lattice)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  tidy=FALSE,
  fig.width=4.5,
  fig.height=3.8,
  out.width=".45\\textwidth",
  out.height=".38\\textwidth",
  comment=NA)

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countries <- c(US = "US", ZA = "ZA")
# List of energy types
energyTypes <- c(Q = "Q", X = "X", U = "U")
@

@@

@@

  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
                 start=(list(lambda=lambdaGuess,alpha=alphaGuess)),
                 data=dataTable)

@@

<<CD Results, results='asis'>>=
print(tablesCD[["US"]], caption.placement="top")
print(tablesCD[["ZA"]], caption.placement="top")
# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCD, caption.placement="top", sanitize.text.function = function(x){x})
@

@@

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
##
cobbDouglasEnergyData <- function(countryName, energyType){

@@

  a <- as.numeric(coef(modelCDe)["a"])
  b <- as.numeric(coef(modelCDe)["b"])
  lambda <- as.numeric(coef(modelCDe)["lambda"])
  alpha <- a
  beta <- b - a
  gamma <- 1.0 - alpha - beta

@@

<<CDq analysis>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countries, cobbDouglasEnergyTable, energyType="Q")
@

<<CDq Results, results='asis'>>=
print(CDqTables[["US"]], caption.placement="top")
print(CDqTables[["ZA"]], caption.placement="top")
# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCDe, sanitize.text.function = function(x){x})

@@

<<CDx analysis>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDxTables <- lapply(countries, cobbDouglasEnergyTable, energyType="X")
@

<<CDx Results, results='asis'>>=
print(CDxTables[["US"]], caption.placement="top")
print(CDxTables[["ZA"]], caption.placement="top")
@

@@

<<CDu analysis>>=
CDuTables <- lapply(countries, cobbDouglasEnergyTable, energyType="U")
@

<<CDu Results, results='asis'>>=
print(CDuTables[["US"]], caption.placement="top")
print(CDuTables[["ZA"]], caption.placement="top")
@

>>>>

Caleb, put your LaTeX code here.

<<setup, include=FALSE>>=
require(knitr)
require(lattice)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  tidy=FALSE,
  fig.width=4,
  fig.height=3.8,
  out.width=".45\\textwidth",
  out.height=".38\\textwidth",
  comment=NA)

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countries <- c(US="US", UK="UK", JP="JP", ZA = "ZA", CN="CN", IR="IR", SA="SA", TZ="TZ", ZM="ZM")
# List of energy types
energyTypes <- c(Q = "Q", X = "X", U = "U")
@

@@

<<creategraph>>=

createCountryGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graph <- xyplot(iGDP ~ iYear, data=dataTable)
  return(graph)
}

#graph <- createCountryGraph("US")
print(createCountryGraph("US"))

@

@@

  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
#                 algorithm = "port",
                 start = list(lambda=lambdaGuess, alpha=alphaGuess),
#                 lower = list(lambda=-Inf, alpha=0),
#                 upper = list(lambda=Inf, alpha=1),
                 data=dataTable)

@@

<<CD Results, results='asis'>>=
print(tablesCD[["US"]], caption.placement="top")
print(tablesCD[["UK"]], caption.placement="top")
print(tablesCD[["JP"]], caption.placement="top")
print(tablesCD[["ZA"]], caption.placement="top")
print(tablesCD[["CN"]], caption.placement="top")
print(tablesCD[["SA"]], caption.placement="top")
print(tablesCD[["IR"]], caption.placement="top")
print(tablesCD[["TZ"]], caption.placement="top")
print(tablesCD[["ZM"]], caption.placement="top")

# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCD, caption.placement="top", sanitize.text.function = function(x){x})
@

@@

<<cobbDouglasEnergy function, include=TRUE, echo=FALSE>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
##
cobbDouglasEnergyData <- function(countryName, energyType){

@@

  a <- as.numeric(coef(modelCDe)["a"])
  b <- as.numeric(coef(modelCDe)["b"])
  print(a)
  print(b)
  lambda <- as.numeric(coef(modelCDe)["lambda"])
  alpha <- a
  beta <- b - a
  gamma <- 1.0 - alpha - beta

@@

<<CDq analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countries, cobbDouglasEnergyTable, energyType="Q")
@

<<CDq Results, results='asis', echo=TRUE, eval=FALSE>>=
print(CDqTables[["US"]], caption.placement="top")
print(CDqTables[["ZA"]], caption.placement="top")
# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCDe, sanitize.text.function = function(x){x})

@@

<<CDx analysis, eval=FALSE>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDxTables <- lapply(countries, cobbDouglasEnergyTable, energyType="X")
@

<<CDx Results, results='asis', eval=FALSE>>=
print(CDxTables[["US"]], caption.placement="top")
print(CDxTables[["ZA"]], caption.placement="top")
@

@@

<<CDu analysis, eval=FALSE>>=
CDuTables <- lapply(countries, cobbDouglasEnergyTable, energyType="U")
@

<<CDu Results, results='asis', eval=FALSE>>=
print(CDuTables[["US"]], caption.placement="top")
print(CDuTables[["ZA"]], caption.placement="top")
@
